% !TEX root = ../main.tex
% The abstract.
% Included by MAIN.TEX

\clearemptydoublepage
\phantomsection
\addcontentsline{toc}{chapter}{Abstract}

\vspace*{2cm}
\begin{center}
{\Large \textbf Abstract}
\end{center}
\vspace{1cm}

Design Optimization is a prospective field of engineering approaches which allows finding solutions for many engineering problems under multiple constraints in an emergence manner\todo{replace:non-analytic?}.
The existing algorithms are computationally expensive and could be sped up by finding approximate results using data-driven methods.
%However, the existing algorithms are computationally expensive and could be sped up by finding heuristic coarse results and using them as an initial educated guess.
In the work, we propose an approach to predicted minimal compliance element material layout using a convolutional neural network model.
Furthermore, we propose novel loss function for the model, as well as pre-processing and post-processing approaches for the data and estimate their influence on the model.
%Model trained on data composed of the existing Topology Optimization cases to quickly infer the approximate results based on the initial conditions. 
The corresponding CNN architecture is proposed and a framework for easy performance analysis and generation of resulting dataset for future use is implemented using Python and Tensorflow.
We analyze the ability of the model to interpolate various parameters and its capacity to be trained\todo{correct?} on small datasets. 
Both 2D and 3D cases are considered and an approach to use model for Topology Optimization performance improvement is proposed and examined.
We analyze possibility to use model-produced\todo{correct?} results as an educated guess for further TO process and estimate the speedup.

%Some more great abstract tips can be found here: 
%\href{https://users.ece.cmu.edu/~koopman/essays/abstract.html}{Great Abstract tips}
